---
title: "Human Activity Recognition"
author: "Christian Sousa"
date: "December 21, 2014"
output: html_document
---

# Summary

The present analysis is based on the _Weight Lifting Exercises Dataset_ obtained from [Ugulino et al](http://groupware.les.inf.puc-rio.br/har) in their research about _Human Activity Recognition_.

The data was collected using several sensors in different parts of the body.
There is a _classe_ column that labels the exercise.

It could be one of the following 5 factors:

- *Class A*: exactly according to the specification
- *Class B*: throwing the elbows to the front
- *Class C*: lifting the dumbbell only halfway
- *Class D*: lowering the dumbbell only halfway
- *Class E*: throwing the hips to the front

The Machine Learning technique selected was *Random Forest* and it could achieve *99.55%* accuracy with a training set consisting of *75%* of the samples.

# Processing

Using the _Caret_ package and loading the data:

```{r, echo = TRUE, cache = TRUE, errors = FALSE, message = FALSE, warning = FALSE}
library(caret)
set.seed(666)

weightLiftingData <- read.csv("pml-training.csv", stringsAsFactors = TRUE)
```

After examinating the _weightLiftingData_ in the R workspace, I discarded several columns that had *NA* or empty values and the ones that didn't seem to add any additional value.

I saved the remaining 52 valid columns in a file *valid_column_names.csv*:

```{r, echo = TRUE, cache = TRUE, errors = FALSE, message = FALSE, warning = FALSE}
columnNames <- read.csv("valid_column_names.csv", stringsAsFactors = FALSE)[[1]]
```

Then using only those columns I generated the training and testing sets:

```{r, echo = TRUE, cache = TRUE, errors = FALSE, message = FALSE, warning = FALSE}
wl <- weightLiftingData[, columnNames]

inTrain <- createDataPartition(wl$classe, p = 0.75, list = FALSE)
training <- wl[inTrain, ]
testing <- wl[-inTrain, ]
```

And execute the training algorithm using _Random Forests_ with a parallel processing setup using 2 cpu cores:

```{r, echo = TRUE, cache = TRUE, errors = FALSE, message = FALSE, warning = FALSE}
library(doMC)
registerDoMC(cores = 2)

tc <- trainControl(method = "cv", number = 3, allowParallel = TRUE)
startTime <- proc.time()
fit <- train(classe ~ ., data = training, method = "rf", trControl = tc, prox = FALSE)
duration <- proc.time() - startTime
```

The training process took

```{r, echo = TRUE, cache = TRUE, errors = FALSE, message = FALSE, warning = FALSE}
duration
```

The evaluation of the _testing_ data and confusion matrix:

```{r, echo = TRUE, cache = TRUE, errors = FALSE, message = FALSE, warning = FALSE}
prediction <- predict(fit, newdata = testing)
confusionMatrix(testing$classe, prediction)
```

With this model and the test cases I can solve the second question:

```{r, echo = TRUE, cache = TRUE, errors = FALSE, message = FALSE, warning = FALSE}
testCasesData <- read.csv("pml-testing.csv", stringsAsFactors = TRUE)
testCases <- testCasesData[, columnNames[-52]]
predict(fit, testCases)
```

